---
title: "Decaying-Dimerization Reaction Set (Gillespie, 2001)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decaying-Dimerization Reaction Set (Gillespie, 2001)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/decaying_dimer.Rmd", output_format = "github_document")
-->

```{r, setseed, echo=FALSE}
set.seed(1)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
```

The Decaying-Dimerization Reaction Set consists of three species and four reaction channels.

```
      S1 --c1--> 0
 S1 + S1 --c2--> S2
      S2 --c3--> S1 + S1
      S2 --c4--> S3
```

Load package
```{r}
library(GillespieSSA)
```

Define parameters
```{r}
parms <- c(c1=1.0, c2=0.002, c3=0.5, c4=0.04)
tf <- 10                                       # Final time
simName <- "Decaying-Dimerizing Reaction Set"  # Name
```

Define initial state vector
```{r}
x0 <- c(s1=10000, s2=0, s3=0)
```

Define state-change matrix
```{r}
nu <- matrix(c(-1, -2, +2,  0,
                0, +1, -1, -1,
                0,  0,  0, +1),
                nrow=3,byrow=TRUE)
```

Define propensity functions
```{r}
a  <- c("c1*s1", "c2*s1*s1", "c3*s2", "c4*s2")
```

Run simulations with the Direct method
```{r direct}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "D",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Explict tau-leap method
```{r etl}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "ETL",
  tau = 0.003,
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Binomial tau-leap method
```{r btl}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "BTL",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Optimized tau-leap method
```{r otl}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "OTL",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```
