---
title: "Kermack-McKendrick SIR model (Brown & Rothery, 1993)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Kermack-McKendrick SIR model (Brown & Rothery, 1993)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<!-- github markdown built using 
rmarkdown::render("vignettes/sir.Rmd", output_format = "github_document")
-->

```{r, setseed, echo=FALSE}
set.seed(1)
knitr::opts_chunk$set(fig.width = 8, fig.height = 6)
```

The Kermack-McKendrick SIR model is defined as
```
dS/dt = -beta*N*S
dI/dt = beta*N*S - gamma*I
dR/dt = gamma*I
```

This model consists of two reactions with the following per capita rates,
```
transmission: beta
recovery:     gamma
```

Load package
```{r}
library(GillespieSSA)
```

Define parameters
```{r}
parms <- c(beta=.001, gamma=.100)
tf <- 100                                      # Final time
simName <- "Kermack-McKendrick SIR"            # Name
```

Define initial state vector
```{r}
x0 <- c(S=500, I=1, R=0)
```

Define state-change matrix
```{r}
nu <- matrix(c(-1,0,1,-1,0,1),nrow=3,byrow=TRUE)
```

Define propensity functions
```{r}
a  <- c("beta*S*I", "gamma*I")
```

Run simulations with the Direct method
```{r direct}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "D",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Explict tau-leap method
```{r etl}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "ETL",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Binomial tau-leap method
```{r btl}
set.seed(2) # for some reason, this does not work with seed = 1
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "BTL",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```

Run simulations with the Optimized tau-leap method
```{r otl}
set.seed(1)
out <- ssa(
  x0 = x0,
  a = a,
  nu = nu,
  parms = parms,
  tf = tf,
  method = "OTL",
  simName = simName,
  verbose = FALSE,
  consoleInterval = 1
) 
ssa.plot(out, show.title = TRUE, show.legend = FALSE)
```
